---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# circletime

<!-- badges: start -->
<!-- badges: end -->

The goal of circletime is to provide tools to model and analyze cyclical biological processes via single-cell RNAseq analysis. To do this, circletime provides a series of tools to generate cyclical pseudotemoral ordering, visualization, and identification of genes associated with progression through the cyclical process.

## Installation

You can install circletime from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("arc85/circletime")
```
## Pseudotemporal ordering of a cyclical process

Here, we demonstrate how to generate a cyclical pseudotemporal ordering and to apply that ordering to our low-dimensional embedded data. We are going to utilize two principal components here for simplicity, but this methodology can be applied to other embeddings that meaningful preserve high-dimensional distances (e.g. diffusion maps or UMAPs).

The single-cell expression data used for this demonstration are derived from [Nestorowa et al, Blood 2016](https://pubmed.ncbi.nlm.nih.gov/27365425/). Usage of this dataset was inspired by the Seurat vignette for regressing out cell cycle effects, linked [here](https://satijalab.org/seurat/v3.2/cell_cycle_vignette.html). Please see also the R files in data-raw (or eventually the pkgdown vignette) for a full description of the pre-processing.

## Order cells through cyclical process

```{r Pseudotemporal-ordering}
library(circletime)
library(ggplot2)

#Visualize embedding
ggplot(cycle.small,aes(x=PC_1,y=PC_3,colour=clusters)) +
	geom_point() +
	theme_bw()

#Calculate pseudotime for all cells
small.pseudotime <- calc_pseudotime(embedding=cycle.small[,c(1,3)],clusters=cycle.small$clusters)

head(small.pseudotime)

#Combine data and visualize pseudotime
cycle.small.pseudotime <- cbind(cycle.small,small.pseudotime)

ggplot(cycle.small.pseudotime,aes(x=PC_1,y=PC_3,colour=pseudotime)) +
	geom_point() +
	theme_bw()

```

## Map progression through pseudotime

It can be useful to visualize the progression through the cyclical process. We demonstrate here how to easily do this using `order_clusters_by_time`.

```{r progression-visualization}

temporal.ordered.clusters <- order_clusters_by_time(embedding=cycle.small.pseudotime[,c(1,3)],clusters=cycle.small.pseudotime$clusters,pseudotime=cycle.small.pseudotime$pseudotime)

#Convert previous clusters to temporally ordered clusters
level.key <- temporal.ordered.clusters$new_cluster_id
names(level.key) <- temporal.ordered.clusters$clusters

cycle.small.pseudotime$ordered.clusters <- dplyr::recode(cycle.small.pseudotime$clusters,!!!level.key)

cycle.small.pseudotime$ordered.clusters <- as.factor(cycle.small.pseudotime$ordered.clusters)

#Visualize progression
ggplot() +
	geom_point(data=cycle.small.pseudotime,aes(x=PC_1,y=PC_3,colour=ordered.clusters)) +
  geom_point(data=temporal.ordered.clusters,aes(x=PC_1_mean,y=PC_3_mean),size=2) +
  geom_path(data=temporal.ordered.clusters,aes(x=PC_1_mean,y=PC_3_mean)) +
	theme_bw()

```

## Identify genes associated with cyclical process

```{r genes-over-cycle}

```
